---
title: "MAST30025 Assingment 3"
author: "Kim Seang CHY"
date: "998008"
output: pdf_document
---


\noindent Question 1:

#Prequestion
```{r}
X = scan(file="assignment3_prob1.txt", what=double())
length(X)
mean(X)
```

\noindent a. Derive the posterior distribution of $\tau$.

Since $x_i|\tau\sim \text{N}(75,\frac{1}{\tau})$ and $\tau \sim\text{Gamma}(2,1)$ Let $X=(x_1,\ldots,x_{100})$.

The likelihood of $X|\tau$ is given by:
\begin{align*}
f(X|\tau)&=\prod_{i=1}^{100}f(x_i|\tau) \\
&\propto (\tau)^{50} \text{exp} \left[\frac{-\tau}{2}\sum_{i=1}^{100}\left(x_i-75 \right)^2 \right]
\end{align*}

Thus $P(\tau|X)$ is given by:
\begin{align*}
P(\tau|X)&\propto f(X|\tau)P(\tau)\\
&\propto (\tau)^{50}\text{exp} \left[\frac{-\tau}{2}\sum_{i=1}^{100}\left(x_i-75 \right)^2 \right] (\tau)\text{exp}(-\tau)\\
&\propto (\tau)^{52-1}\text{exp} \left[-\tau(1+\frac{1}{2}\sum_{i=1}^{100}(x_i-75)^2)\right]
\end{align*}

By inspection we can see $\tau|X$ is gamma distribution with $\alpha=52$ and $\beta=1+\frac{1}{2}\sum_{i=1}^{100}(x_i-75)^2$ or $\tau|X \sim \text{Gamma}(52,1+\frac{1}{2}\sum_{i=1}^{100}(x_i-75)^2)$.


```{r}
#Calculating therotical beta value
beta = 0.5*sum((X-75)^2)+1
beta
alpha = 2+0.5*length(X)
alpha
```
Thus $\tau|X \sim \text{Gamma}(52,1805.65)$.


\noindent b. Derive the posterior predictive distribution for new score $\tilde x$, $P(\tilde x|x_1,\ldots,x_{100})$.

From q1 part a we know $\tau|X \sim \text{Gamma}(52,1805.65)=\text{Gamma}(\alpha,\beta)$ and $\tilde x|\tau \sim \text{N}(75,\frac{1}{\tau})$, where $X=(x_1,\ldots,x_{100})$.

\begin{align*}
P(\tilde x|x_1,\ldots,x_{100})&=\int_{0}^{\infty}P(\tilde x|\tau)P(\tau|X) \text{ d}\tau \\
&= \int_{0}^{\infty} \sqrt{\frac{\tau}{2\pi}}\text{exp}\left[\frac{-\tau}{2}(\tilde x - 75)^2 \right] \frac{\beta^{\alpha}}{\Gamma(\alpha)}\tau^{\alpha-1}\text{exp}(-\beta \tau) \text{ d}\tau \\
&= \frac{\beta^{\alpha}}{\Gamma(\alpha)\sqrt{2\pi}}\int_{0}^{\infty}\text{exp}\left[-\tau\left(\frac{(\tilde x - 75)^2}{2}+\beta \right)\right] \tau^{\frac{2\alpha+1}{2}-1} \text{ d}\tau \\
&= \frac{\beta^{\alpha}}{\Gamma(\alpha)\sqrt{2\pi}} \frac{\Gamma(\frac{2\alpha+1}{2})}{\left[\frac{\left(\tilde x-75 \right)^2}{2} +\beta \right]^{\frac{2\alpha+1}{2}}}
\end{align*}

Using Gamma distribution normalising constant to solve integration.

\begin{align*}
P(\tilde x|x_1,\ldots,x_{100})&=\frac{\Gamma(\frac{2\alpha+1}{2})}{\Gamma(\frac{\alpha}{2})\sqrt{2\pi \beta}}\left[\frac{(\tilde x - 75)^2}{2\beta}+1\right]^{-\frac{-2\alpha+1}{2}} \\
&= \frac{\Gamma(\frac{2\alpha+1}{2})}{\Gamma(\frac{2\alpha}{2})\sqrt{2\alpha \pi \frac{\beta}{\alpha}}}\left[\frac{1}{2\alpha}\frac{(\tilde x - 75)^2}{\frac{\beta}{\alpha}}+1\right]^{-\frac{-2\alpha+1}{2}}
\end{align*}


```{r}
v=alpha*2
v
a=75
a
b=beta/alpha
b
```

By inspection we can see the predictive distribution for new $\tilde x|(x_1,\ldots,x_{100})\sim \text{t}(104,75,34.72)$, which is a student-t distribution with three parameter of 104, 75, 34.72.

\pagebreak

\noindent Question 2: 


Using the information given we will simulate gamma distribution with parameter $\alpha\geq1$ and $\beta$, as stated below

Let
\begin{align*}
f(x)=\frac{h(x)^{\alpha-1}e^{-h(x)}h'(x)}{\Gamma(\alpha)} \\
d=\alpha-1/3, \\
c=\frac{1}{\sqrt{9d}},\\
h(x)=d(1+cx)^3\\
g(x) = d \log((1+c*x)^3)-d*(1+c*x)^3+d
\end{align*}


Using $f(x)*\Gamma(\alpha) \propto \text{exp}(g(x))$, we can simulate $f(x)$ by simulating $\text{exp}(g(x))$ using the general rejection method with $p(x)=e^{\frac{-x^2}{2}}$ as an envelope with $X\sim N(0,1)$ with either $X \geq \frac{-1}{c}$ or $h(x) \geq0$.\


The algorithm used follow the step below:
\begin{enumerate}
  \item Simulate X from $N(0,1)$.
  \item If $X \geq \frac{-1}{c}$ return X else start from step 1
  \item Generate $Y\sim U(0,p(X))$
  \item If $Y < \text{exp(g(X))}$ return$\frac{\text{exp}(g(X))}{\beta}$ else return to step 1
  \item Repeat until we get the number of n sample required
\end{enumerate}
\

```{r}
#Simulating Gamma Sample with alpha = a, beta = b, number of required sample = n
sim.gamma <- function(a,b,n){
  gamma.sample = c()
  for (i in 1:n){
    gamma.sample[i] = X.Gamma(a)/b
  }
  return(gamma.sample)
}


#Function to Stimulate X and apply general rejection method to g(x) with p(x) as
#an envelope function. Step 1, 2, 3 and Part of 4
X.Gamma <- function(a){
  d=a-1/3
  c=1/sqrt(9*d)
  x.sample=c()
  h <- function(x) d*(1+c*x)^3
  while(TRUE){
    while(TRUE){
      x.point = rnorm(1)
      if (x.point >= -1/c) break
    }
    Y <- runif(1,0,p(x.point))
    if (Y<exp.g(x.point,a)) return(h(x.point)) 
  }
}

#Let p(x)=exp(-x^2/2)
p <-function(x){
  x=exp(-x^2/2)
  return(x)
}


#Define exp(g(x))
exp.g <- function(x,a){
  d=a-1/3
  c=1/sqrt(9*d)
  q <- function(x) (1+c*x)^3
  g <- function(x) d*log(q(x))-d*q(x)+d
  x= exp(g(x))
  return(x)
}

#Running the simulation and Plotting qq-plot
g <- sim.gamma(1.2,3,1000)
plot(qgamma(1:1000/1001, 1.2, 3), sort(g))
abline(0, 1, col="red")

```
\
Looking at the qq-plot we can see that our simulation are close to the pseudo-random of Gamma(1.2,3) for 1000 samples.



