---
title: "Assignment 1"
author: "Name: Kim S CHY"
date: "Student ID:998008"
output: pdf_document
---

\noindent 1. Fit a binomial regression model to the O-rings data from the Challenger disaster, using a probit link. You must use R (but without using the glm function); I want you to work from first principles.

\noindent (a) Compute MLEs (maximum likelihood estimates) of the parameters in the model.
```{r}
#Loading Data
library(faraway)
data("orings")
str(orings)

#Define MLE_Function
MLE_f <- function(beta, orings) {
  eta <- cbind(1, orings$temp) %*% beta
  p <- pnorm(eta)
  return(sum( orings$damage*log(p/(1-p)) +6*log(1-p)))
}
#Calculating Beta.hat
(beta_hat <- optim(c(10,-.1),MLE_f, orings=orings, control = 
                     list(fnscale=-1))$par)
beta_hat
```

\noindent (b) Compute 95% CIs for the estimates of the parameters. You should show how you derived the Fisher information.\

\noindent Let $p_i=F(\eta_i)=\int_{-\infty}^{\eta_i}\frac{e^{\frac{-t^2}{2}}}{\sqrt{2\pi}}dt$, $f(t)=\frac{e^{\frac{-t^2}{2}}}{\sqrt{2\pi}}$ and $\eta_i=\beta_0+\beta_1x_i$.

The log-likelihood function is given by:
\begin{align*}
l(\beta_{0},\beta_{1})&=c+\sum_{i=1} \left[y_i \log (p_i)+(m_i-y_i) \log (1-p_i) \right]\\
& =c+\sum_{i=1} \left[y_i \log (F(\eta_i))+(m_i-y_i) \log (1-F(\eta_i)) \right]
\end{align*}

\begin{align*}
\frac{dl}{d\beta_0}=\sum_{i=1} \left[y_i \frac{f(\eta_i)}{F(\eta_i)}+(m_i-y_i)\frac{f(\eta_i)}{F(\eta_i)-1} \right]
\end{align*}

\noindent Let $u=\eta_i^2=(\beta_0+\beta_1x_i)^2 \implies \frac{df}{du}=\frac{-2e^{\frac{-u}{2}}}{\sqrt{2\pi}}$ and $\frac{du}{d\beta_0}=2\eta_i$\

\noindent From the above we get $\frac{df}{d\beta_0}=-\eta_if(\eta_i)$; hence

\begin{align*}
\frac{dl^2}{d\beta_0^2}&=\sum_{i=1} \left[y_i \left( \frac{-\eta_if(\eta_i)}{F(\eta_i)}-\frac{(f(\eta_i))^2}{(F(\eta_i))^2}    \right)+(m_i-y_i) \left(\frac{-\eta_if(\eta_i)}{F(\eta_i)-1}-\frac{(f(\eta_i))^2}{(F(\eta_i)-1)^2}  \right) \right]\\
&=\sum_{i=1} \left[y_i \left(\frac{-\eta_if(\eta_i)}{F(\eta_i)}+\frac{\eta_if(\eta_i)}{F(\eta_i)-1} -\frac{(f(\eta_i))^2}{(F(\eta_i))^2}+\frac{(f(\eta_i)^2)}{(F(\eta_i)-1)^2} \right) -m_i\left(\frac{\eta_if(\eta_i)}{F(\eta_i)-1}+\frac{(f(\eta_i))^2}{(F(\eta_i)-1)^2}\right)\right]\\
&= \sum_{i} \left[ y_i f(\eta_i)\left(\frac{f(\eta_i)(2F(\eta_i)-1)+\eta_iF(\eta_i)(F(\eta_i)-1)}{(F(\eta_i))^2(F(\eta_i)-1)^2} \right)-m_i\left(\frac{\eta_if(\eta_i)}{F(\eta_i)-1}+\frac{(f(\eta_i))^2}{(F(\eta_i)-1)^2}\right)\right]
\end{align*}\

\noindent Since $Y_i\sim \text{bin}(m_i,p_i)\implies E(Y_i)=p_im_i$; hence:
\begin{align*}
\mathbb{E}\left(-\frac{d^2l}{d\beta_0^2}\right)&=\sum_{i} -\left[ m_i f(\eta_i)\left(\frac{f(\eta_i)(2F(\eta_i)-1)+\eta_iF(\eta_i)(F(\eta_i)-1)}{(F(\eta_i))(F(\eta_i)-1)^2} \right)-m_i\left(\frac{\eta_if(\eta_i)}{F(\eta_i)-1}+\frac{(f(\eta_i))^2}{(F(\eta_i)-1)^2}\right)\right]\\
&= \sum_{i=1} \left[ \frac{m_i*f(\eta_i)}{F(\eta_i)(F(\eta)-1)}\right]\\
&=\sum_{i=1} \left[2m_i(f(\eta_i))^2\left(\frac{1}{p_i-p_i^2} \right)     \right]
\end{align*}

\begin{align*}
\frac{dl^2}{d\beta_0d\beta_1}=\sum_{i=1} \left[ x_iy_i f(\eta_i)\left(\frac{f(\eta_i)(2F(\eta_i)-1)+\eta_iF(\eta_i)(F(\eta_i)-1)}{(F(\eta_i))^2(F(\eta_i)-1)^2} \right)-m_i\left(\frac{\eta_if(\eta_i)}{F(\eta_i)-1}+\frac{(f(\eta_i))^2}{(F(\eta_i)-1)^2}\right)\right]
\end{align*}

\begin{align*}
\mathbb{E}\left(-\frac{d^2l}{d\beta_0d\beta_1}\right)&= \sum_{i=1}- \left[x_im_i(f(\eta_i))^2\left(\frac{1-F(\eta_i)}{F(\eta_i)(F(\eta_i)-1)^2} \right)     \right]\\
&= \sum_{i=1} \left[x_im_i(f(\eta_i))^2\left(\frac{1}{p_i-p_i^2} \right)    \right] 
\end{align*}

\begin{align*}
\frac{dl^2}{d\beta_1^2}=\sum_{i} \left[ (x_i)^2y_i f(\eta_i)\left(\frac{f(\eta_i)(2F(\eta_i)-1)+\eta_iF(\eta_i)(F(\eta_i)-1)}{(F(\eta_i))^2(F(\eta_i)-1)^2} \right)-m_i\left(\frac{\eta_if(\eta_i)}{F(\eta_i)-1}+\frac{(f(\eta_i))^2}{(F(\eta_i)-1)^2}\right)\right]
\end{align*}


\begin{align*}
\mathbb{E}\left(-\frac{d^2l}{d\beta_0d\beta_1}\right)&= \sum_{i=1}- \left[(x_i)^2m_i(f(\eta_i))^2\left(\frac{1-F(\eta_i)}{F(\eta_i)(F(\eta_i)-1)^2} \right)     \right]\\
&=\sum_{i=1} \left[(x_i)^2m_i(f(\eta_i))^2\left(\frac{1}{p_i-p_i^2} \right) \right]
\end{align*}




The fisher information is given by:
\[
\mathcal{I}(\beta)=
\begin{bmatrix}
\sum_{i=1} \left[m_i(f(\eta_i))^2\left(\frac{1}{p_i-p_i^2} \right)     \right] & \sum_{i=1} \left[x_im_i(f(\eta_i))^2\left(\frac{1}{p_i-p_i^2} \right)    \right] \\
\sum_{i=1} \left[x_im_i(f(\eta_i))^2\left(\frac{1}{p_i-p_i^2} \right)     \right] & \sum_{i=1} \left[(x_i)^2m_i(f(\eta_i))^2\left(\frac{1}{p_i-p_i^2} \right)   \right]
\end{bmatrix}
\]

\noindent Finding standard error for parameter
```{r}
f <- function(t){
  exp(-(t^2)/2)/(sqrt(2*pi))
}

phat <- pnorm(beta_hat[1]+orings$temp*beta_hat[2])
h <- function(t){
  1/(t-t^2)
}
I11 <- sum(6*(f(phat)^2)*h(phat))
I12 <- sum(6*orings$temp*(f(phat))^2*h(phat))
I22 <- sum(6*(orings$temp^2)*(f(phat))^2*h(phat))
Iinv <- solve(matrix(c(I11, I12, I12, I22), 2,2))
sqrt(Iinv[1,1])
sqrt(Iinv[2,2])
```


\noindent The 95% Confidence Interval for $\beta_0$ and $\beta_1$ is:

```{r}
#Beta 0
c(beta_hat[1]-qnorm(0.025)*sqrt(Iinv[1,1]),beta_hat[1]+qnorm(0.025)*sqrt(Iinv[1,1]))

#Beta 1

c(beta_hat[2]-qnorm(0.025)*sqrt(Iinv[2,2]),beta_hat[2]+qnorm(0.025)*sqrt(Iinv[2,2]))

```


\noindent c. Perform a likelihood ratio test for the significance of the temperature coefficient.


```{r}
MaxlogL.F <- MLE_f(beta_hat,orings)
MaxlogL.F
y <- orings$damage 
n <- rep(6, length(y))
phatN <- sum(y)/sum(n) 
MaxlogL.R = sum(orings$damage)*log(phatN) + sum(6-orings$damage)*log(1-phatN)
MaxlogL.R

```

\noindent LR-Test statistic and p-value for LR-test
```{r}
LR = -2*(MaxlogL.R - MaxlogL.F)
LR
pchisq(LR, df=1,lower=FALSE)
```


\noindent d. Compute an estimate of the probability of damage when the temperature equals 31 Fahrenheit (your estimate should come with a 95% CI, as all good estimates do).\



```{r}
#Computing Prediction for 31 degree
phat <- pnorm(beta_hat[1]+beta_hat[2]*31)
phat

#Computing 95% CI 31 degree
si2 <- matrix(c(1, 31), 1, 2) %*% Iinv %*% matrix(c(1, 31), 2, 1) 
etahat = beta_hat[1] + beta_hat[2]*31 
eta_l = etahat - 2*sqrt(si2) 
eta_r = etahat + 2*sqrt(si2)

c(pnorm(eta_l),pnorm(eta_r))
```


\noindent e. Make a plot comparing the fitted probit model to the fitted logit model. To obtain the fitted logit model, you are allowed to use the glm function.

 

```{r}
#Building Model using GLM
probitmodel <-  glm(cbind(damage,6-damage) ~ temp, family=binomial(link="probit"), orings) 
logitmodel <- probitmodel <-  glm(cbind(damage,6-damage) ~ temp, family=binomial, orings)
temp_rise <- seq(25, 85, 1)

#Getting sequence of response
logitresp <- predict(logitmodel,list(temp=temp_rise),type="response")
probitresp <- predict(probitmodel,list(temp=temp_rise), type = "response")

#Plotting
plot(temp_rise,logitresp, type = "l", col="red",xlab = "Temperature in Fahrenheit",
     ylab="Probability of Damage")
lines(temp_rise, probitresp, col="Blue")
```

\noindent We can see that the line perfectly overlap one another making it indifference between using either the probit or logit model.


\pagebreak
\noindent Q2.\

\noindent a.Please estimate the amount of increase in the $\log$(odds) when the bmi increases by 7.


\noindent Loading Data and Building Model
```{r}
#Loading Data
data()
missing <- with(pima, missing <- glucose==0 | diastolic==0 | triceps==0 | bmi == 0)
pima_subset = pima[!missing, c(6,9)]
str(pima_subset)


#Building Model using GLM
```


```{r}
modelQ2 <- glm(cbind(test, 1-test)~bmi, family=binomial, data=pima_subset)
summary(modelQ2)
```

\noindent Please Estimate the amount increase
```{r}
x <- predict(modelQ2, newdata = list(bmi=7),type="link", se.fit=TRUE)

log_odd <- function(phat){
  exp(phat/(1-phat))
}

log_odd(x$fit)
```


\noindent b. Compute a 95% CI for the estimate. \


```{r}
log_odd(c(x$fit-2*x$se.fit, x$fit+2*x$se.fit))
```


\pagebreak
\noindent 3.The gamma distribution with shape $\nu >0$ and rate $\lambda>0$ has pdf:

\begin{align*}
f(x;\nu,\lambda)=\frac{\lambda^{\nu}}{\Gamma(\nu)}x^{\nu-1}e^{-\lambda x}
\end{align*}
\noindent for $x>0$.\

\noindent a.Show that the gamma distribution is an exponential family.\
\noindent Want to show that $f(x;\nu,\lambda)=\frac{\lambda^{\nu}}{\Gamma(\nu)}x^{\nu-1}e^{-\lambda x}$ can be written as $f(x;\theta,\phi)=\text{exp}\left[ \frac{x\theta-b(\theta)}{a(\phi)}-c(x,\phi)\right]$.\

\begin{align*}
f(x;\nu,\lambda)&=\frac{\lambda^{\nu}}{\Gamma(\nu)}x^{\nu-1}e^{-\lambda x}\\
&=\text{exp} \left[\nu \log(\lambda)-\log(\Gamma(\nu))+(\nu-1)\log(x)-\lambda x\right]\\
&=\text{exp} \left[-(\lambda x - \nu \log(\lambda))+(\nu-1) \log(x)-log(\Gamma(\nu))  \right]\\
&=\text{exp} \left[\frac{\frac{\lambda x}{\nu}-\log(\lambda)}{\frac{-1}{\nu}}+(\nu-1)\log(x)-\log(\Gamma(\nu)) \right]
\end{align*}

\noindent Let $\theta=\frac{\lambda}{\nu}$ and $\phi=\frac{1}{\nu}$.\
\noindent Hence $\lambda=\theta\nu=\frac{\theta}{\phi} \implies \log(\lambda)=\log(\theta)-\log(\phi)$.

The above become:
\begin{align*}
f(x;\theta,\phi)&=\text{exp} \left[ \frac{\theta x-\log(\theta)}{-\phi}+\frac{\log(\phi)}{\phi} +\left(\frac{1}{\phi}-1 \log(x)-\log(\Gamma(\frac{1}{\phi})) \right)\right]
\end{align*}

Now Let $b(\theta)=\log(\theta)$; $a(\phi)=-\phi$ and $c(x,\phi)=\left(\frac{1}{\phi}-1 \log(x)-\log(\Gamma(\frac{1}{\phi})) \right)$. Thus: 
\begin{align*}
f(x;\theta,\phi)=\text{exp}\left[ \frac{x\theta-b(\theta)}{a(\phi)}-c(x,\phi)\right]
\end{align*}
\noindent Hence it is part of the exponential family.\

\noindent b. Obtain the canonical link and the variance function.


\noindent Since, $\mathbb{E}(X)=b'(\theta)=\frac{db}{d\theta}=\theta^{-1}=\frac{\nu}{\lambda}$ and the canonical link function is given by $\left[ b'(\mu)\right]^{-1}$.\
This implied the canical link function is $g(\mu)=\theta=\left[ b'(\mu)\right]^{-1}=\mu^{-1}$.\

\noindent Since $b''(\theta)=\frac{d^2b}{d\theta^2}=-\theta^{-2}$ and the variance function is given by: $v(\mu)=-\theta^{-2}\cdot \mu^{-1}=-\mu$.Thus the variance function is given by:

\begin{align*}
\text{Var}  X=v(\mu)a(\phi)=\mu\phi=\frac{\mu}{\nu}
\end{align*}



