---
title: ""
author: ""
date: ""
output:
  pdf_document: default
---

\textbf{Question 2:}


Pre-question
```{r}
#Setting up Matrix Y
Y <- matrix(c(22,23,24,22,26,16,18,19,28,27,29,29),12,1)
Y

#Setting up matrix X
X <- matrix(rep(0,36),12,4)
X[,1] = 1 ;X[1:5,2]=1; X[6:8,3]=1; X[9:12,4]=1
X

```

\vspace{0.2cm}

\textbf{a.}Find a conditional inverse for $X^TX$, using the algorithm given in Theorem 6.2.


```{r}
#Computing X^tX
XtX <- t(X)%*%(X)
XtXc <- matrix(rep(0,36),4,4)
#Computing one of the conditional inverse
XtXc[2:4,2:4] <-solve(XtX[2:4,2:4])
XtXc
```

\vspace{0.9cm}
\textbf{b.}Find $s^2$.

```{r}
#Computing s2
df <- 12-3 #n=12 and reading X we can see r(X)=3 
I <- diag(12)
SSres <- t(Y)%*%(I-X%*%XtXc%*%t(X))%*%Y
SSres
s2 <- SSres/df
```

\vspace{1cm}
\textbf{c.}Is $\mu + 2\tau_1 + \tau_2$ estimable?
```{r}
#Testing for estimablility using Thm 6.10
t1 <- matrix(c(1,2,1,0),4,1)
t(t1)%*%XtXc%*%XtX
```

\vspace{0.5cm}

\noindent Since $\textbf{t}_1^T(X^TX)^C(X^TX) \neq \textbf{t}_1^T$, by theorem 6.10, $\mu + 2\tau_1 + \tau_2$ is not estimable.



\vspace{0.5cm}

\textbf{c.}Find a 90\% confidence interval for the lifetime of the 2nd type of bulb.
```{r}
#Computing Beta
b <- XtXc%*%t(X)%*%Y
b
#Computing 90% CI interval for 2nd type of bulb
t2 <- matrix(c(0,0,1,0),4,1)
tstat_halfalpha <- qt(0.95,df)
t2.std <- tstat_halfalpha*sqrt(s2*(t(t2)%*%XtXc%*%t2))
CI.90 <- c(t(t2)%*%b - t2.std,t(t2)%*%b + t2.std)
CI.90
```






\textbf{d.}Test the hypothesis that there is no difference in lifetime between 1st and 3rd types of bulb, at the 5\% significance level.\

\noindent Testing for $H_0:\tau_1=\tau_3 \text{ vs } H_1:\tau_1\neq \tau_3$ 
```{r}
#Testing for H_0:tau.1=tau.3 and H_1:tau.1!=tau.3 
C <- matrix(c(0,1,0,-1),1,4)
#C has a rank of 2
dst <- 0
numerator <- (t(C%*%b-dst)%*%solve(C%*%XtXc%*%t(C))%*%C%*%b-dst)
Fstat <- (numerator/2)/s2
pf(Fstat,2,df,lower=F)
```

\vspace{0.5cm}
We can firmly reject the null at a 5\% statistical significant, thus there are difference in lifetime between 1st and and 3rd types of bulb.


\pagebreak


\textbf{Question 4:}\


\textbf{a.}Plot the data, using different colours and/or symbols for male and female records. Without drawing diagnostic plots, do you think that this data satisfies the assumptions of the linear model? Why or why not?\

```{r}
library(ggplot2)
mile2 <- read.csv(file="mile2.csv")
mile2$Gender.f <- factor(mile2$Gender)
g1 <- ggplot(mile2, aes(x=Year, y=Time,colour=Gender)) + geom_point()
g2 <- g1 + labs(x="Year", y="Time in second",
         title="World Record Time for the one mile Run 
From 1861 to 1999 for Male \n From 1967 to 1996 for Female")
g2 + theme_bw()
```

\vspace{0.5cm}
From the graph, we can see a linear relationship for Male. Homoscedasticity seem to also be met  as beside the big outlier of 1861, we can the point seem be randomly distributed around the regression.For the big outlier we will need to do further analysis to see if the point has a high crook's distance.
For female beside the lack of a big outlier, the same thing could be said about the linear relationship and homoscedasticity assumption.

\pagebreak

\textbf{b.}Test the hypothesis that there is no interaction between the two predictor variables. Interpret the result in the context of the study.
```{r}
#Determine if there are interaction using R
with(mile2, interaction.plot(Year,Gender.f, Time))
model <- lm(Time~Gender.f+Year, mile2)
imodel <- lm(Time~Gender.f*Year, mile2)
anova(model,imodel)
```

Looking at the interaction we can see that there an interaction between year and gender as the line is not parallel to one another. Furthermore, this is confirmed by our statistical test where we reject the null at 5\%. Thus indicating

\pagebreak

\textbf{c.}Write down the final fitted models for the male and female records. Add lines corresponding to these models to your plot from part (a).

```{r}
library(MASS)
library(Matrix)
g3 <- g2 + geom_smooth(method = lm)
g3+theme_bw()
#Computing Beta
Y <- mile2$Time
n <- length(Y)
X <- matrix(0,n,6)
X[,1] <- 1
X[cbind(1:n,as.numeric(mile2$Gender.f)+1)] <-1
X[,4] <- mile2$Year
X[cbind(1:n,as.numeric(mile2$Gender.f)+4)] <- mile2$Year
XtX <- t(X)%*%X
r <- rankMatrix(X)[1]
XtXc <- matrix(0,6,6)
M <- XtX[c(2:3,5:6),c(2:3,5:6)]
XtXc[c(2:3,5:6),c(2:3,5:6)] <-  solve(M)
b <- XtXc%*%t(X)%*%Y
b
```

The model is given by: 
\begin{align*}
\text{Time}=0.0321-2309.4247\text{ Female}-953.747\text{ Male}+0.0322-1.0337\text{ Female:Year}-0.3662\text{ Male:Year}
\end{align*}\

Where Female is an indicator taking the value of 1 for female and 0 for male. Similarly, Male is also an indicator with value of 1 for male and 0 for female.

Hence the respective subpopulation model for female is given by:
\begin{align*}
\text{Time}_{Female}=2309.4247-1.0337\text{ Year}
\end{align*}

For male is given by:
\begin{align*}
\text{Time}_{Male}=953.747-0.3662\text{ Year}
\end{align*}

\textbf{d.}Calculate a point estimate for the year when the female world record will equal the male world record. Do you expect this estimate to be accurate? Why or why not?\


```{r}
(1355.6777866)/(0.6675093)
#Computing time taken in 2030.95 for female
t1 <- matrix(c(1,1,0,1,(1355.6777866)/(0.6675093),0),6,1)
t(t1)%*%b
#Computing time taken in 2030.95 for female
t2 <- matrix(c(1,0,1,1,0,(1355.6777866)/(0.6675093)),6,1)
t(t2)%*%b
```

Letting $\text{Time}_{Female}=\text{Time}_{Male}$ or $2309.4247-1.0337\text{ Year}=953.747-0.3662\text{ Year}$, we get the estimate year where male and female will be equal to each other near the end of the year 2030 with a time of 210.1045 second. \
First of all such a point would not be estimable on our model.Secondly, our model may not have the power to predict what happen in the future because the model assumed perpuity improvement instead of taking into account of the physical limitation gap between Male and Female.\

\textbf{e.}Is the year when the female world record will equal the male world record an estimable quantity? Is your answer consistent with part (d)?

```{r}
t(t1)%*%XtXc%*%XtX

t(t2)%*%XtXc%*%XtX
```

Using theorem 6.10 we can see that both $\textbf{t}_1^T(X^TX)^CX^TX) \neq\textbf{t}_1$. Similarly, $\textbf{t}_2^T(X^TX)^CX^TX) \neq\textbf{t}_2$. Thus, such a point estimate is not estimatable which was consistent with the answer given in part c.\

\textbf{f.}Calculate a 95\% confidence interval for the amount by which the gap between the male and female world records narrow every year.


```{r}
confint(imodel)
```

The Gender.fMale:Year is the differences between female and male record each year. Reading from the above we get the 95\% CI for amount of gap by which female and male record narrow every year of $(0.462, 0.873)$ second every year.


\textbf{g.}Test the hypothesis that the male world record decreases by 0.4 seconds each year.

We want to test $H_0:\beta_2=-0.4$
```{r}
C <- matrix(c(0,0,0,0,0,1),1,6)
#C has a rank of 1
dst <- -0.4
numerator <- t(C%*%b-dst)%*%solve(C%*%XtXc%*%t(C))%*%(C%*%b-dst)
Fstat <- (numerator)/s2
Fstat
pf(Fstat,1,df,lower=F)
```

We can reject the null that the male world record will decreases by -0.4 second at 5\% statistical significant.


\pagebreak


\textbf{Question 5:}You wish to perform a study to compare 2 medical treatments (and a placebo) for a disease. Treatment 1 is an experimental new treatment, and costs \$5000 per person. Treatment 2 is a standard treatment, and costs \$2000 per person. Treatment 3 is a placebo, and costs \$1000 per person. You are given \$100,000 to complete the study. You wish to test if the treatments are effective, i.e., $H_0: \tau_1 = \tau_2 = \tau_3$.

\textbf{a.}Determine the optimal allocation of the number of units to assign to each treatment.\

We are testing for $H_0: \tau_1 = \tau_2 = \tau_3$ or equivalently we are testing the contrast differences $\tau_1 - \tau_3$ and $\tau_2 - \tau_3$. By doing this I put a heavy emphasis on placebo group relative to other two groups. \

The first constraint we have is $5n_1+2n_2+n_3=100$
\begin{align*}
\text{var} ( \widehat{ \tau_i - \tau_3} )= \sigma^2  \left( \frac{1}{n_i} + \frac{1}{n_3} \right)\\
\\
f(n_1,n_2,n_3,\lambda)= \sigma^2 \left( \frac{1}{n_1} +\frac{1}{n_2} + \frac{2}{n_3} \right) + \lambda \left( \sum_{i=1}^3 n_i -n  \right)
\end{align*}

Solve $\frac{df}{dn_i}=0$, we get $n_3=\sqrt{2}n_2=\sqrt{2}n_2$. Subbing this back into our first constraint of $5n_1+2n_2+n_3=100$. 


```{r}
#Solving for n3
100/(1+(7/sqrt(2)))

#Solving for n1=n2
16.80744/sqrt(2)
```


We get $n_1=n_2=12$ and $n_3=16$ after rounding out the samples.

\textbf{b.}Perform the random allocation. You must use R for randomisation and include your R commands and output.
```{r}
#Performing random selection
n <- c(12,12,16)
nsum <- sum(n)
x <- sample(nsum, nsum)
n1 <- x[1:n[1]]
n2 <- x[n[1]+1:n[2]]
n3 <- x[n[2]+1:n[3]]
```

Thus our assignment to treatment 1 for $y_i$ will be:
```{r}
n1
```


Our assignment to treatment 2 for $y_i$ will be:
```{r}
n2
```


And our assignment to placeobo for $y_i$ will be:
```{r}
n3
```






